---
title: 'Data Wrangling Part 1: Introduction to dplyr'
author: Marlon Schumacher
date: '2018-11-10'
slug: data-wrangling-part-1-introduction-to-dplyr
categories:
  - R
  - Tutorial
tags:
  - dplyr
  - data wrangling
  - data manipulation
image:
  caption: ''
  focal_point: ''
output: 
  blogdown::html_page:
    toc: false
---

# Introduction to dplyr

<p align="justify">When it comes to data manipulation, `dplyr` is a very powerful package with very intuitive functions. There are six important functions you will use often, if it comes to data manipulation:</p>

+ `filter()`
+ `select()`
+ `mutate()`
+ `arrange()`
+ `group_by()`
+ `summarise()`

<p align="justify">In addition to these functions `dplyr` offers you the ability to use the pipe-operator out of the `magrittr` package. This operator is very useful and I highly recommend you to get used to it. Except these functions, `dplyr` offers a lot more functions which are very useful. And, like always, you should look into the [reference manual of dplyr](https://cran.r-project.org/web/packages/dplyr/dplyr.pdf) for more details! I will mostly use the data frame `gapminder` for explaining the functions.</p>

# filter()

<p align="justify">n many cases you need to filter your dataset because you are only interested in specific cases or you want to exclude some cases. For that, the `filter()` function provides an easy way to do that.</p>

Let's take a look into the `gapminder` df, which is an excerpt of the Gapminder data on life expectancy, GDP per capita, and population by country:
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(gapminder)
knitr::kable(head(gapminder))
```

Now, different aspects of the data may be of interest:

+ data for specific year(s)
+ data for specific continent(s)
+ data for specific continent(s) and year(s)

A very simple filtering could look like this one:
```{r}
# filtering by the year 1997
gapminder %>% 
  filter(year == 1997) %>% 
  head() %>% 
  kableExtra::kable()
```

<p align="justify">It is also possible to filter with more than one condition. If we are only interested in the year 1997 and the continent asia, we could use the following code.</p>
```{r}
# filtering by the year 1997 and the continent asia
gapminder %>% 
  filter(year == 1997 & continent == "Asia") %>% 
  head() %>% 
  kableExtra::kable()
```

<p align="justify">As you can imagine, there are many logical operators that you can use. I guess, most of you know these operators already.</p>

+ `<`: less than
+ `<=`: less than or equal to
+ `==`: equal to
+ `!=`: not equal to
+ `>=`: greater than or eqal to
+ `>`: greater than
+ `%in%`: matching argument
+ `!x`: not x (e.g. often used for NA exclusion `!is.na(variable)`)

The following operators can be used to combine different conditions:

+ `&`: filtering by condition 1 **and** contiditon 2
+ `|`: filtering by condition 1 **or** condition 2

You will see a few of these operators in action later.

# select() function

<p align="justify">The `select()` function is a very simple one but can be useful, if the data frame is very large but you only need a few variables. You can select specific coloumns and you can also *delete* sepcific coloumns.</p>

```{r}
# let's select the coloumns country, continent, year and lifeExp
gapminder %>% 
  select(country, continent, year, lifeExp) %>% 
  head() %>% 
  kableExtra::kable()

# a shorter way to select these coloumns
gapminder %>% 
  select(country:lifeExp) %>% 
  head() %>% 
  kableExtra::kable()
```

If we want to delete only a few columns of the df, we can do this by using `-`:
```{r}
# deleting the column population
gapminder %>% 
  select(-pop) %>% 
  head() %>% 
  kableExtra::kable()
```

# mutate() function

<p align="justify">You can create new columns with the `mutate()` function. Moreover, it is possible to use the `mutate()` function in combination with other functions (e.g. `case_when()`). In the following code I create a new variable, where the value is 10% of the population size.</p>

```{r}
# creating a new coloumn where the value is 10% of the population
gapminder %>% 
  mutate(new_pop = round(pop*0.1)) %>% 
  head() %>% 
  kableExtra::kable()
```

<p align="justify">Now, let's create a new variable with three categories. Therefore, we will use the `mutate()` function and the `case_when()` function in combination. In the following example the new variable should include three categories: low life expectancy (lifeExp < 50), medium life expectancy (lifeExp >= 50 & <70) and high life expectancy (>= 70).</p>

```{r}
# creating three categories
gapminder %>% 
  mutate(lifeExp_cat = case_when(
    lifeExp < 50 ~ "low",
    lifeExp < 70 & lifeExp >= 50 ~ "medium", 
    lifeExp >= 70 ~ "high"
  )) %>% 
  head() %>% 
  kableExtra::kable()
```

<p align="justify">As you can see, you can add a bunch of conditions with the `case_when()` function. Moreover, with this combination of these functions it is very easy to manipulate specific cases if it's needed. Personally, I use the combination of these functions very often. In this way, it is very easy to implement specific encodings.</p>

# group_by() & summarise() function

<p align="justify">If you take a look into the data for the first time, it is often very useful to make some exploratory data analysis. The `group_by()` function in combination with the `summarise()` function makes it easy. For example, we can group by continent, calculate the mean for the life expectancy and plot the results for each continent over the time. I did that all with the pipe operator (`%>%`). If you are unfamiliar with the use of this operator, you should familiarize yourself with it. In my opinion, you can work much more efficiently with it.</p>

```{r}
library(ggplot2)
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(mean_lifeExp = mean(lifeExp)) %>% 
  ggplot(aes(x = year, y = mean_lifeExp, col = continent)) +
    geom_line()
```


