---
title: 'Data Wrangling Part 1: Introduction to dplyr'
author: Marlon Schumacher
date: '2018-11-16'
slug: data-wrangling-part-1-introduction-to-dplyr
categories:
  - R
  - Tutorial
tags:
  - dplyr
  - Tidyverse
  - data wrangling
  - data manipulation
image:
  caption: ''
  focal_point: ''
output: 
  blogdown::html_page:
    toc: true
---

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(rmdformats)
```

# Introduction to dplyr

<p align="justify">When it comes to data manipulation, `dplyr` is a very powerful package with very intuitive functions. There are six important functions you will use often, if it comes to data manipulation:</p>

+ `filter()`
+ `select()`
+ `mutate()`
+ `arrange()`
+ `group_by()`
+ `summarise()`

<p align="justify">In addition to these functions `dplyr` offers you the ability to use the pipe-operator out of the `magrittr` package. This operator is very useful and I highly recommend you to get used to it. Except these functions, `dplyr` offers a lot more functions which are very useful. And, like always, you should look into the [reference manual of dplyr](https://cran.r-project.org/web/packages/dplyr/dplyr.pdf) for more details! I will mostly use the data frame `gapminder` for explaining the functions.</p>

# filter()

<p align="justify">In many cases you need to filter your dataset because you are only interested in specific cases or you want to exclude some cases. For that, the `filter()` function provides an easy way to do that.</p>

Let's take a look into the `gapminder` df, which is an excerpt of the Gapminder data on life expectancy, GDP per capita, and population by country:
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(gapminder)
knitr::kable(head(gapminder))
```

Now, different aspects of the data may be of interest:

+ data for specific year(s)
+ data for specific continent(s)
+ data for specific continent(s) and year(s)

A very simple filtering could look like this one:
```{r}
# filtering by the year 1997
gapminder %>% 
  filter(year == 1997) %>% 
  head() %>% 
  kableExtra::kable()
```

<p align="justify">It is also possible to filter with more than one condition. If we are only interested in the year 1997 **and** the continent asia, we could use the following code.</p>
```{r}
# filtering by the year 1997 and the continent asia
gapminder %>% 
  filter(year == 1997 & continent == "Asia") %>% 
  head() %>% 
  kableExtra::kable()
```

<p align="justify">As you can imagine, there are many logical operators that you can use. I guess, most of you know these operators already.</p>

+ `<`: less than
+ `<=`: less than or equal to
+ `==`: equal to
+ `!=`: not equal to
+ `>=`: greater than or eqal to
+ `>`: greater than
+ `%in%`: matching argument
+ `!x`: not x (e.g. often used for NA exclusion `!is.na(variable)`)

The following operators can be used to combine different conditions:

+ `&`: filtering by condition 1 **and** contiditon 2
+ `|`: filtering by condition 1 **or** condition 2

You will see a few of these operators in action later.

# select() function

<p align="justify">The `select()` function is a very simple one but can be useful, if the data frame is very large but you only need a few variables. You can select specific coloumns and you can also delete sepcific coloumns.</p>

```{r}
# let's select the coloumns country, continent, year and lifeExp
gapminder %>% 
  select(country, continent, year, lifeExp) %>% 
  head() %>% 
  kableExtra::kable()

# a shorter way to select these coloumns
gapminder %>% 
  select(country:lifeExp) %>% 
  head() %>% 
  kableExtra::kable()
```

If we want to delete only a few columns of the df, we can do this by using `-`:
```{r}
# deleting the column population
gapminder %>% 
  select(-pop) %>% 
  head() %>% 
  kableExtra::kable()
```

# mutate() function

<p align="justify">You can create new columns with the `mutate()` function. Moreover, it is possible to use the `mutate()` function in combination with other functions (e.g. `case_when()`). In the following code I create a new variable, where the value is the log of the population size.</p>

```{r}
gapminder %>% 
  mutate(log_pop = log(pop)) %>% 
  head() %>% 
  kableExtra::kable()
```

<p align="justify">Now, let's create a new variable with three categories. Therefore, we will use the `mutate()` function and the `case_when()` function in combination. In the following example the new variable should include three categories: low life expectancy (lifeExp < 50), medium life expectancy (lifeExp >= 50 & <70) and high life expectancy (>= 70).</p>

```{r}
gapminder %>% 
  mutate(lifeExp_cat = case_when(
    lifeExp < 50 ~ "low",
    lifeExp < 70 & lifeExp >= 50 ~ "medium", 
    lifeExp >= 70 ~ "high"
  )) %>% 
  head() %>% 
  kableExtra::kable()
```

<p align="justify">As you can see, you can add a bunch of conditions with the `case_when()` function. Moreover, with this combination of these functions it is very easy to manipulate specific cases if it's needed. Personally, I use the combination of these functions very often. In this way, it is very easy to implement specific encodings.</p>

# arrange() function

<p align="justify">This function can be used for sorting the data by one variable. You can also use multiple variables for this function. It is good for a first exploratory analysis or if you want to familiarize yourself with the data. The following code sorts the data by `lifeExp`:</p>

```{r}
gapminder %>% 
  arrange(lifeExp) %>% 
  head() %>% 
  kableExtra::kable()
```

<p align="justify">If you want to sort your data in descending order instead of ascending order, you can do this by using `desc()`:</p>
```{r}
gapminder %>% 
  arrange(desc(lifeExp)) %>% 
  head() %>% 
  kableExtra::kable()
```

<p align="justify">As mentioned above, it is possible to sort the data by multiple variables. The following code sorts the data by year in ascending order and then by lifeExp in descending order:</p>

```{r}
gapminder %>% 
  arrange(year, desc(lifeExp)) %>% 
  head() %>% 
  kableExtra::kable()
```


# group_by() & summarise() function

<p align="justify">If you take a look into the data for the first time, it is often very useful to make some exploratory data analysis. The `group_by()` function in combination with the `summarise()` function makes it easy. For example, we can group by continent, calculate the mean for the life expectancy and plot the results for each continent over the time. I did that all with the help of the pipe operator (`%>%`). If you are unfamiliar with the use of this operator, you should familiarize yourself with it. In my opinion, you can work much more efficiently with it. In short, the `%>%`-operator can be seen as a concatenation of functions.</p>

```{r}
library(ggplot2)
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(mean_lifeExp = mean(lifeExp)) %>% 
  ggplot(aes(x = year, y = mean_lifeExp, col = continent)) +
    geom_line()
```

<p align="justify">Of course you do not have to plot immediately. However, you can easily extract important data from your dataset. Let's group again by continent but now we will calculate the mean of life expectancy, the total population size and the median of the GDP for each continent:</p>

```{r}
gapminder %>% 
  filter(year == 1997) %>% 
  group_by(continent) %>% 
  summarise(mean_lifeExp = mean(lifeExp),
            total_pop = sum(as.numeric(pop)),
            median_gdp = median(gdpPercap)) %>% 
  head() %>% 
  kableExtra::kable()
```

