---
title: "Structural Equation Modeling in R"
author: "Marlon Schumacher"
date: '2018-10-24'
slug: structural-equation-modeling-in-r
categories:
  - R
  - Tutorial
tags:
  - SEM
  - Structural Equation Modeling
  - lavaan
  - semplot
  - fit indices
image:
  caption: ''
  focal_point: ''
bibliography: bibliography_001.bib 
csl: mcstatistics.csl
output: 
  blogdown::html_page:
    toc: true
---

## lavaan: A Package for SEM
<p align="justify">In my pervious post I made a [short introduction to Structural Equation Modeling (SEM)](https://www.mcstatistics.eu/post/structural-equation-modeling-a-short-introduction/). But now it's about using R for SEM. However, to use R for SEM we need the package `lavaan`. For the visualization of the models we also need the package `semPlot`. As you probably know already the installation of the packages is very simple:</p>

```{r eval = FALSE}
install.packages("lavaan")
install.packages("semPlot")
```

## Syntax for Modelling
The modely syntax of lavaan is used for specifying a SEM. There are three operators for model specification:

+ `=~` is used to specify latent variables/factors
+ `~~` is used to specify (residual) (co)variances
+ `~`  is used to specify rgeressions

An example model specification might look like this:

```{r}
example <- "
# measurement model
F1 =~ 1*V1 + V2 + V3
F2 =~ 1*V4 + V5 + V6

# (residual) (co)variances
V1 ~~ V1 # variance
V1 ~~ V3 # residual correlation

# regressions
F1 ~ F2"
```

## Model Fitting
<p align="justify">Now, let's specify the first model. I will use an own dataset from a previous work. It's also possible to use the built-in datasets `HolzingerSwineford1939` or `PolitcialDemocracy` from lavaan to become familiar with the package. Moreover, the functions `cfa()` and `sem()` are currently similar but they may differ in the future. It's also possible to use one of the robust estimators, if you're fitting the model.</p>

```{r message=FALSE, warning=FALSE, include=FALSE}
load("allbus_imp.RData")
library(magrittr)
library(tidyverse)
library(lavaan)
data <- allbus_imp %>% 
  select(-age, -sex, -work, -mc04)

colnames(data) <- c("X1", "X2", "X3", "X4", "X5", "X6", "Y1", "Y2", "Y3", "Y4", "weight")

model <- "
# measurement model
F1 =~ 1*X1 + X2 + X3 + X4 + X5 + X6
F2 =~ 1*Y1 + Y2 + Y3 + Y4

# regression
F1 ~ F2"

# fitting the model
fit <- cfa(model, data = data)

# summary of the fit with standardized values and fit measurements
summary(fit, standardized = T, fit.measures = T)
```


```{r message=FALSE, warning=FALSE}
library(lavaan)
model <- "
# measurement model
F1 =~ 1*X1 + X2 + X3 + X4 + X5 + X6
F2 =~ 1*Y1 + Y2 + Y3 + Y4

# regression
F1 ~ F2"

# fitting the model
fit <- cfa(model, 
           data = data,
           estimator = "MLM") # robust estimator 

# summary of the fit with standardized values and fit measurements
summary(fit, standardized = T, fit.measures = T)
```

<p align="justify">If you want to select specific fit measurements, you can use the function `fitmeasures()`. By using the command `fit.measures = "all"` you will get all available fit measures.</p>

```{r message=FALSE, warning=FALSE}
fitmeasures(fit, fit.measures = c("chisq", "cfi", "rmsea", "srmr"))
```

<p align="justify">Apart from the fit indices, the standardized residuals can also be used to interpret the model fit. Unfortunately there is no function to get the standardized residuals. However, a function can be written very fast for this. Using `cov2cor` on the covariance matrices gives us the fitted and observed correlation matrices. To get the residual correlation matrix we need to subtracting the fitted correlation matrix from the observed correlation matrix.</p>

```{r}
std_residuals <- function(fit) {
  cov <- list(observed = inspect(fit, "sampstat")$cov,
              fitted = fitted(fit)$cov)
  cor <- list(observed = cov2cor(cov$observed),
              fitted = cov2cor(cov$fitted))
  cor$residual <- cor$observed - cor$fitted
lapply(cor, function(x) round(x, 2))
}

std_residuals(fit)
```


## (Standardized) Estimated Parameter Change
<p align="justify">To identify potential model misspecification, you can use `modificationindices()`. If the SEPC is $\geq.20$, there may be a possible misspecification. However, adjustments to the model should be based on a theoretical justification.</p>

```{r message=FALSE, warning=FALSE}
# first 10 rows as output & sorting by mi-value
head(modificationindices(fit, sort. = T), 10) %>% 
  dplyr::select(-sepc.lv, -sepc.nox) %>% 
  kableExtra::kable()
```

## Visualization
<p align="justify">For visualization of an structural equation model the package `semPlot` offers many possibilities. I would recommend you to look in the reference manual of this package. There are a lot of specifications you can do. However, the specification that I'm using should cover everything necessary for the beginning. In general, an adjustment of the rotation and the layout will probably be necessary.</p>

```{r message=FALSE, warning=FALSE}
library(semPlot)
semPaths(fit, # fitted model
         nCharNodes = 0, # no abbreviation in node labels
         whatLabels = "std", # standardized parameter estimate 
         residuals = F, # excluding residuals and variances
         sizeLat = 10, # width latent
         sizeLat2 = 10, # height latent
         sizeMan = 6, # width manifest
         edge.label.cex = 0.90, # font size of parameters
         layout = "tree2", # type of layout
         rotation = 2) # rotation of the layout
```

<p align="justify">In addition, the following visualization is often used when it comes to using the function `semPaths()`. There is a weighting of the edges depending on the standardized parameter estimates.</p>

```{r message=FALSE, warning=FALSE}
semPaths(fit, # fitted model
         nCharNodes = 0, # no abbreviation in node labels
         what = "std", # standardized parameter estimates as weighted edges 
         whatLabels = "std", # standardized parameter estimate 
         residuals = F, # excluding residuals and variances
         sizeLat = 10, # width latent
         sizeLat2 = 10, # height latent
         sizeMan = 6, # width manifest
         edge.label.cex = 0.90, # font size of parameters
         layout = "tree2", # type of layout
         rotation = 2) # rotation of the layout
```

## Sampling Weights
<p align="justify">With lavaan version 0.6-3 there are two ways for weighting. On the one hand, you can use the package `lavaan.survey`for taking sampling weights into account. On the other hand, it's possible to specify the sampling weights within the `cfa()` function.</p>

```{r message=FALSE, warning=FALSE}
fit_2 <- cfa(model, 
           data = data,
           estimator = "MLM")

library(lavaan.survey)
# specify survey design
sd_fit <- survey::svydesign(id = ~1,
                  weights = ~weight,
                  data = data)

# fitting the model again while taking the survey design into account
surveyfit <- lavaan.survey(lavaan.fit = fit_2,
                           survey.design = sd_fit)

summary(surveyfit, standardized = T)
```

```{r}
# weighting within the cfa() function
weighted_fit <- cfa(model, 
           data = data,
           estimator = "MLM",
           sampling.weights = "weight") # sepcify sampling weights

summary(weighted_fit, standardized = T)
```


