---
title: "Working with APIs: Cryptocurrencies"
author: Marlon Schumacher
date: '2019-04-21'
slug: working-with-apis
categories:
  - R
tags:
  - API
  - R
  - Forecasting
  - json
image:
  caption: ''
  focal_point: ''
bibliography: bibliography_001.bib 
csl: mcstatistics.csl
output: 
  blogdown::html_page:
    toc: true
---

## Introduction 

Working with APIs is a really powerful way to get new data you can work with. Especially if you want to make some personal projects by yourself or with friends/colleagues. However, to get full access to the data with an API, you will need to pay for it in most cases. I would like to show you a very simple way to work with APIs. 

The packages `tidyverse`,` httr` and `jsonlite` are the basic packages I mostly use when I am working with APIs. If you are working with APIs you will notice that you will often need to handle lists. The package `rlist` can be useful for handling lists.  But in most cases the `dplyr` package is sufficient. Since I work with time series here, I also use the `anytime` package next to `lubridate`.

```{r}
pacman::p_load(tidyverse, httr, jsonlite, rlist, coindeskr, anytime, forecast, tseries, predtoolsTS)
```


## API for Cryptocurrencies & Getting the Data

There are some different APIs you can use to get the price for different cryptocurrencies. However, you often need to pay for the APIs to get a key with access to all historical data. With the [API from cryptocompare](https://min-api.cryptocompare.com) you are able to get all historical data in day format. You can find the documentation of the API [here](https://min-api.cryptocompare.com/documentation). First of all, we need an API key for using the API from cryptocompare. Most APIs require a key which you can get for free mostly. Moreover, the urls for APIs are mostly divided into two parts: the base url and the endpoint. If we take the API from cryptocompare, we have the base url `https://min-api.cryptocompare.com/data`. Depending on the data you want to get, you need to use a different endpoint for the url. Since wie want to get the historic data for specific currencies in day format, we need to use the endpoint `/histoday`. Now, we only need the paramters to get the specific data we want. You can find the parameters in the documentation of the API you're working with.

If we want to get the price for Bitcoin for the past 365 days, we can use the following code.
```{r include=FALSE}
key_cc <- "0f30ada169bfc36bc21d5dd0cfc7ebc8d21846a449565ae8ca3bb382444d66f9"
```

```{r}
url_hist <- "https://min-api.cryptocompare.com/data/histoday"

btc_data <- GET(url_hist, query = list(api_key = key_cc, # providing the api key (as a string)
                                       fsym = "BTC", # define the cryptocurrency
                                       tsym = "USD",
                                       limit = 365)) # defining the number of days (depending on the time format you're working with)

btc_data
```

As you can see, we made a successful call. The status code needs to be 200. Otherwise, something went wrong with your call (e.g. wrong API key). However, the object we're getting with the `GET()` function out of the `httr` package, is not a simple data frame, it's a JASON format. Therefore, we need to transform this object into a data frame. To do this, we will use the `content()` function from `httr` and transform the list into a data frame. 

```{r}
btc_df_365 <- content(btc_data, as = "parse")$Data %>% 
  bind_rows() %>% 
  mutate(date = anytime::anytime(time)) %>% 
  mutate(Date = as.Date(date, format = "Y%-m%-d%"), # transforming the date format into ISO...IMPORTANIMPORTANT
         ID = "BTC", # adding the column ID
         asset = "Bitcoin") # adding the column asset

head(btc_df_365) %>% 
  kableExtra::kable()
```

We got the data we wanted. However, we would need to write a code for each cryptocurrency that we want. To avoid a long code, we can simply write a function that needs a vector with the cryptocurrencies and stores the data in a data frame. Moreover, we can set the parameter `allData` to "true" for getting every day.

```{r}
# writing a little function
get_crypto_history <- function(asset = "BTC", url = FALSE, api_key = FALSE){
  
  # creating little dummy df which is used to combine all data frames
  df <- data.frame(time = 1, close = 1.0, high = 1.0, low = 1.0, open = 1.0, volumefrom = 1.0, 
                   volumeto = 1.0, date = as.POSIXct("2010-07-17 02:00:00"), Date = as.Date("2001-01-01"))
  
  # for loop to get the data for all given cryptos
  for(i in seq_along(asset)){
    df_raw <- GET(url, query = list(api_key = api_key,
                                    fsym = asset[i],
                                    tsym = "USD",
                                    allData = "true")) # need to be a string
    
      df_raw <- content(df_raw, as = "parse")$Data %>% 
      bind_rows() %>% 
      mutate(date = anytime::anytime(time)) %>% 
      mutate(Date = as.Date(date, format = "Y%-m%-d%"),
             ID = asset[i])
      
    df <- bind_rows(df, df_raw)
  }
  df <- df[-1,]
  df
}

# using the written function for getting the data from 4 coins
coin_df <- get_crypto_history(asset = c("BTC", "BCH", "BTG", "ETH"), 
                              url = url_hist, 
                              api_key = key_cc) %>% 
  mutate(asset = case_when(
    ID == "BTC" ~ "Bitcoin",
    ID == "BCH" ~ "Bitcoin Cash",
    ID == "BTG" ~ "Bitcoin Gold",
    ID == "ETH" ~ "Ethereum"))

head(coin_df) %>% 
  kableExtra::kable()
```

## Plotting the Data

```{r}
coin_df %>% 
  ggplot(aes(x = Date, y = close, col = asset)) +
  geom_line() +
  ggtitle("Bitcoin, Bitcoin Hard Forks & Ethereum", subtitle = "Price (USD) since 17/07/2010") +
  xlab("") +
  ylab("") +
  theme_bw() +
  scale_x_date(breaks = scales::date_breaks("6 months"),
               date_labels = "%b %y") +
  labs(color="Cryptocurrency") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle=45, vjust = 0.5))
```


```{r}
coin_df %>% 
  ggplot(aes(x = Date, y = close, col = asset)) +
  geom_line() +
  ggtitle("Bitcoin, Bitcoin Hard Forks & Ethereum", subtitle = "Price (log10) since 17/07/2010") +
  xlab("") +
  ylab("") +
  scale_y_log10() +
  theme_bw() +
  scale_x_date(breaks = scales::date_breaks("6 months"),
               date_labels = "%b %y") +
  labs(color="Cryptocurrency") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle=45, vjust = 0.5)) 
```

